Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Prediction in the Swiss Alps

### 2017-01-23 Optimization case study - JoH/corrections/5 - English editing.tex
Elsevier english editing.

%TODO: read again
%TODO: Elsevier english editing
%TODO: relative improvement of the crpss...
%TODO: homogenize hours: 6 a.m. or 06 h ??
%TODO: follow author guide

@@

\begin{abstract}
Analogue methods provide a statistical precipitation prediction based on synoptic predictors supplied by general circulation models or numerical weather prediction models. The method samples a selection of days in the archives that are similar to the target day to predict, and consider the set of their corresponding observed precipitation (the predictand) as the conditional distribution for the target day. The relationship between predictors and predictands relies on some parameters that characterize how and where the similitude between two atmospheric situations is defined.

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations: (i) it cannot automatically choose the pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters, and (iii) it cannot easily handle new degrees of freedom. In this work, a global optimization approach relying on genetic algorithms could optimize all parameters jointly and automatically. It allowed taking into account parameters inter-dependencies, and selecting objectively some parameters that were manually selected beforehand, which e.g. obviates the need of assessing a high number of combinations of the pressure levels and the temporal windows of the predictor variables.

The global optimization was applied to some variants of the analogue method for the Rh\^{o}ne catchment in the Swiss Alps. The performance scores increased compared to reference methods, and this even to a greater extent for days with high precipitation totals. The resulting parameters were found to be relevant and coherent between the different subregions of the catchment. Moreover, they were obtained automatically and objectively, which reduces efforts invested in exploration attempts when adapting the method to a new region or for a new predictand. In addition, the approach allowed for new degrees of freedom, such as a possible weighting between the pressure levels, and non overlapping spatial windows.
\end{abstract}

@@

The analogue method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969} that similar situations in terms of the atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Horton2016a, Hamill2015, BenDaoud2016} or a climate downscaling context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

In real-time forecasting, it is used mainly by practitioners, notably hydropower companies or flood forecasting services. They need to anticipate water yields or to issue early flood warnings several days in advance. The classical forecasting chain consists in using limited area models (e.g. AROME, or COSMO) forced by global NWP (numerical weather prediction) models with a lower resolution. However, their use require very important processing capacities, and the resulting forecast still present large uncertainties and biases. Although these outputs are essential, they can be supplemented by other sources of forecasts providing useful information. In contrast to local NWP models, AMs can transform at low cost the synoptic-scale information provided by the global NWP model into precipitation forecasts, by using the natural local behavior in response to synoptic-scale influence stored in the archive of observed precipitation. Running an AM approach is so quick that it can search for analogues for every day up to ten days ahead, eventually for the different traces of an ensemble forecast and/or issued by different NWP models (e.g. NOAA-GFS or ECMWF-IFS), in a matter of minutes.

In climate studies AMs are used to downscale the outputs of general circulation model (GCM) or regional climate models (RCMs) simulation runs \citep{Dayon2015} or to reconstruct past weather conditions \citep{Caillouet2016}. In future climate studies, RCMs are often used to dynamically downscale precipitation to a local scale. However, even though the relevance of RCMs'outputs increases, a bias correction of the outputs is still often required, particularly in complex terrain. Moreover, their application is computer-intensive, which makes it difficult to cover all combinations of climate scenarios and GCMs. Therefore, the idea is to bypass the small scale simulations and to go from the large scale situation to the end variables like precipitation by statistical downscaling \citep{Maraun2010}.

@@

The method can be designed with multiple successive subsampling steps, or analogy levels, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criterion itself, and finally the number of analogue situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep{Bontron2004, Radanovics2013}, i.e. by optimizing each single parameter one at a time, in an arbitrarily chosen order, with no or little reassessment. This sequential approach has therefore strong limitations: (i) it cannot automatically choose the optimal pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it cannot easily handle new degrees of freedom, such as a possible weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictors combinations (variables, pressure levels, temporal windows), and presents a high risk of ending in a local optimum due to subjective initial choices and lack of consideration of parameter inter-dependencies. Other calibration methods exist for specific applications, such as radar images \citep{Panziera2011, Foresti2015a}.

With the perspective to overcome these limitations, a global optimization by genetic algorithms (GAs) was introduced. An intensive assessment work resulted in recommendations of parametrization of GAs in order to optimize AMs successfully \citep{Horton2016}. The present paper is based on these recommendations and apply them for precipitation prediction for the upper Rh\^{o}ne catchment in the Swiss Alps, using AMs of varying complexity. It aims at illustrating the relevance of a fully-automatic, objective, and global, optimization technique for AMs. The applications are indeed numerous, as AMs have to be adapted to every new location they are applied, or to any new predictand they should predict.

Data, AMs, and the optimization techniques (sequential and GAs) are presented in section \ref{sec:data_methods}. The results are first given for the optimization of the analogy of the atmospheric circulation only (section \ref{sec:optim_circul}), before being extended to a method adding a second level of analogy on moisture variables (section \ref{sec:optim_moisture}). General discussions (section \ref{sec:discussion}) and conclusions (section \ref{sec:conclusions}) follow.

@@

The study area is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l.\ and the area is 5524~km$^{2}$. This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project that aimed at real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. Even though the region is rather small, the meteorological influences related to extreme weather conditions vary substantially within it \citep[see][]{Horton2012}. Indeed, a high spatial variability of the precipitation climatology exists, which is due to the complex orography of the region, and the mix of various meteorological influences. Based on different climatological analyses, the precipitation gauging stations in the catchment were clustered in ten subregions (Fig.\ \ref{fig:map}) :

@@

AMs rely on two types of data: predictors, that are atmospheric variables describing the state of the atmosphere at a synoptic scale, and the predictand, which is the local weather variable one wants to predict.

@@

The predictand (which is to be predicted) is here the daily precipitation (6~a.m. to 6~a.m. the next day) measured at the MeteoSwiss' stations network, for the period 1961-2008. The time series from every available gauging station were averaged over the ten subregions (Fig.\ \ref{fig:map}) of approximately 500~km$^{2}$ each in order to smooth local effects \citep{Obled2002, Marty2012}. It helps accounting for local variability, mainly when convective processes are involved, which increases slightly the prediction skill.

It must be stressed that the predictand is here sa temporally cumulated variable, compared to the meteorological predictors, which may be considered instantaneous. Depending on the duration of the accumulation period (here 24~h, but could have been 6, 12~h, or more than 24~h), the choice of predictors will vary.

The 48 years precipitation dataset was divided into a calibration period (CP) and a validation period (VP). Using data independent from the CP to validate the results is very important in order to assess the robustness of the proposed improvements and to avoid over-parametrization of the method.

In order to reduce potential bias related to trends linked to climate change or to the evolution in measurement techniques, the selection of the VP was evenly distributed over the entire series \citep{BenDaoud2010}. Thus, one year every six years was selected for validation, which represents a total of 8 years for the VP and 40 for CP. The choice of the sequence was made in order to have similar statistical characteristics between the CP and the VP.

\subsection{The analogue method}
\label{sec:references}

Multiple variations of the analogue method exist, and most of them will not be detailed hereafter \cite[see][for more comprehensive listings]{BenDaoud2016}. However, there are mainly two parameterizations that are most often used for precipitation prediction and that were considered as reference: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

The method based on the analogy of the synoptic circulation consists in the following steps (Table \ref{table:params_R1}): the similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criterion \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window:

\begin{equation}
\label{eq:S1}
S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the difference in geopotential height between the \textit{i}th pair of adjacent points of the gridded data describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both North and East directions over the selected spatial domain. The smaller the S1 values, the more similar the pressure fields.

\citet{Bontron2005} show that the geopotential height at 500~hPa (Z500) and 1000~hPa (Z1000) are the best first predictors of the NCEP/NCAR reanalysis I dataset, and that the S1 criterion performs better than scores based on absolute distances. The reason for such better results is that the S1 criterion allows comparing the circulation patterns, by means of the gradients, rather than the absolute value of the geopotential height, which represent better the flow direction. To cope with seasonal effects, candidate dates are extracted within a period of four months centered around the target date, for every year of the archive. This method using two geopotential heights will be named 2Z.

The $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day. The number of analogues, $N_{1}$, is a parameter to calibrate. Then, the daily observed precipitation amount of the $N_{1}$ resulting dates provide the empirical conditional distribution considered as the probabilistic prediction for the target day.

The other most known parametrization adds a second level of analogy on moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirms that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis I considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled within the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When this second level of analogy is added, a higher number of analogues $N_{1}$ is kept on the first level. Moisture fields are not as well predicted by NWP models as pressure variables. It implies that the 2Z-2MI method, when used in real-time forecasting, is very dependent on the skill of the NWP model in predicting moisture fields, and thus its use is often restricted to the first lead times.

@@

The performance assessment in the present context consists in verifying the prediction of an ensemble-probabilistic technique. The set of precipitations values collected with each analogue can be considered as a sample drawn from the conditional distribution associated with the current circulation. The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $n$ can be written:

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{n} \sum_{i=1}^{n} \left(  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right) 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value 1 otherwise. The mean CRPS is averaged on the calibration, respectively the validation periods, on all days.

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution of precipitation from the entire archive as the reference. The CRPSS (Continuous Ranked Probability Skill Score) is thus defined as following:

@@

Note however that the choice of the reference does not matter so much when assessing potential improvements of the method, since we shall consider more its relative increase or decrease rather than the CRPSS absolute value.

@@

AMs are usually calibrated by a semi-automatic sequential procedure elaborated by \citet{Bontron2004} \cite[see also ][]{Radanovics2013, BenDaoud2016}. The calibration technique optimizes the spatial windows on which the predictors are compared and the number of analogues, for every level of analogy, by maximizing the performance score (CRPSS). However, the different analogy levels are calibrated sequentially, and the meteorological variables, the pressure levels and the temporal windows are chosen manually. The procedure, as defined by \citet{Bontron2004}, consists of the following steps:

\begin{enumerate}
	\item Manual selection of the following parameters:
	\begin{enumerate}
		\item meteorological variable,
		\item pressure level,
		\item temporal window (hour of the day),
		\item initial analogue numbers.
	\end{enumerate}

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification, for the analogy level considered, of the most skilled unitary cell of all predictors jointly, over a large domain, by a full scanning of the grid.
		\item From this most skilled cell, the spatial window is expanded by successive iterations in the direction of greater performance gain until no improvement is reached.
		\item The number of analogue situations $N_{1}$ is then reconsidered and optimized for the current level of analogy.
	\end{enumerate}
	\item A new level of analogy can then be added, based on other variables (such as the moisture index) with chosen pressure levels, temporal windows, and initial number of analogues $N_{2}$. The procedure starts again from step 2 (calibration of the spatial window and the number of analogues) for the new level. The parameters calibrated on the previous analogy levels are fixed and do not change. 
	\item Finally, the numbers of analogues $N_{1}$ and $N_{2}$ for the different levels of analogy are reassessed by systematic increments.
\end{enumerate}

@@

Genetic algorithms (GAs) were developed by \citet{Holland1992b} and \citet{Goldberg1989}. They are part of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, which get inspiration from some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs seek the global optimum on a complex surface, theoretically without restriction, which is of interest for AMs, that are characterized by a complex high-dimensional error function having multiple local optima. Practically, GAs allow rapidly approaching satisfactory solutions, but they do not provide the optimum solution for sure \citep{Zitzler2004a}. It is indeed mainly a matter of time. When the optimizer gets closer to the global optimum, any new improvement takes more time to appear (see for example the slow-down of the improvements over generations in Figure \ref{fig:evolution}), and the final adjustment of the parameters can be very time consuming \citep{Back1993a}. For problems that require a significant amount of time in order to evaluate the objective function, as in the case of AMs (because it needs to make a prediction for every day of the CP), the number of generations has to be limited in order to get reasonable processing time. Thus, different acceptable solutions can result from one or more optimization runs \citep{Holland1992b}. This is both a strength and a weakness of GAs: they are very good at exploring complex parameter spaces in order to identify the most promising areas, but they will not necessarily always find the best solution with the optimal values of all parameters \citep{Holland1992b}.

The optimizations were here performed based on the recommended GAs parametrization for AMs as described in \citet{Horton2016}. As the optimization is mostly sensitive to the mutation operator (that randomly changes some values in the parameters sets), parallel optimizations are considered with variants of this operator, according to \citet{Horton2016}:

\begin{itemize}
	\item 3x non-uniform mutation \citep{Michalewicz1996} with varying parameters,
	\item 1x multi-scale mutation \citep{Horton2016},
	\item 2x chromosome of adaptive search radius \citep{Horton2016}
\end{itemize}

A population size of 500 individuals (i.e. parameter sets of the AM to be detailed hereunder) was considered, and the optimization was stopped when the best individual (with the highest CRPSS performance score) did not evolve for 20 generations (cycles of the optimization).

\section{Optimization of the circulation analogy}
\label{sec:optim_circul}

The analogy of the atmospheric circulation was optimized for the ten subregions (Section \ref{sec:case_study})  independently. We started from the most simple AM, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. The detailed results of the intermediate stages are not provided in this paper \citep[see][for the details]{Horton2012a}.

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose simulatneously the number of analogues, both spatial windows with no overlapping constraint (i.e. they can differ from one pressure level to another), as well as the temporal windows (hours of observation of the geopotential), which cannot be achieved with the sequential calibration technique. The performance score (CRPSS) was slightly improved, with these limited degrees of freedom, in regards to the 2Z reference method calibrated with the sequential procedure. Some tests showed that most of the gains were due to the non-overlapping spatial windows. It demonstrated that the optimizer was able to get relevant parameters for a simple method.

>>>>

%TODO: read again
%TODO: Elsevier english editing
%TODO: relative improvement of the crpss...
%TODO: homogenize hours: 6 a.m. or 06 h ??

@@

\begin{abstract}
Analogue methods provide a statistical precipitation prediction based on synoptic predictors supplied by general circulation models or numerical weather prediction models. The method samples a selection of days in the archives that are similar to the target day to be predicted, and consider their set of corresponding observed precipitation (the predictand) as the conditional distribution for the target day. The relationship between the predictors and predictands relies on some parameters that characterize how and where the similarity between two atmospheric situations is defined.

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations: (i) it cannot automatically choose the pressure levels and temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between parameters, and (iii) it cannot easily handle new degrees of freedom. In this work, a global optimization approach relying on genetic algorithms was able to optimize all parameters jointly and automatically. It allowed consideration of parameter inter-dependencies, and objective selection of some parameters that were manually selected beforehand, which, among other things, obviates the need to assess a large number of combinations of pressure levels and temporal windows of predictor variables.

The global optimization was applied to some variants of the analogue method for the Rh\^{o}ne catchment in the Swiss Alps. The performance scores increased compared to reference methods, especially for days with high precipitation totals. The resulting parameters were found to be relevant and coherent between the different subregions of the catchment. Moreover, they were obtained automatically and objectively, which reduces the effort that needs to be invested in exploration attempts when adapting the method to a new region or for a new predictand. In addition, the approach allowed for new degrees of freedom, such as a possible weighting between pressure levels, and non-overlapping spatial windows.
\end{abstract}

@@

The analogue method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969} that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Horton2016a, Hamill2015, BenDaoud2016} or a climate downscaling context \citep[e.g. ][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

In real-time forecasting, it is used mainly by practitioners, notably hydropower companies or flood forecasting services, that need to anticipate water yields or issue early flood warnings several days in advance. The classical forecasting chain consists of using limited area models (e.g. AROME, or COSMO) forced by global NWP (numerical weather prediction) models with a lower resolution. However, their use requires very important processing capacities, and the resulting forecast still presents large uncertainties and biases. Although these outputs are essential, they can be supplemented by other sources of forecasts providing useful information. In contrast to local NWP models, AMs can transform at low cost the synoptic-scale information provided by the global NWP model into precipitation forecasts, by using the natural local behaviour in response to synoptic-scale influences stored in the archive of observed precipitation. Running an AM approach is fast enough that it can search for analogues for each day, up to ten days ahead, eventually for the different traces of an ensemble forecast and/or those issued by different NWP models (e.g. NOAA-GFS or ECMWF-IFS), in a matter of minutes.

In climate studies, AMs are used to downscale the outputs of a general circulation model (GCM) or regional climate model (RCM) simulation runs \citep{Dayon2015} or to reconstruct past weather conditions \citep{Caillouet2016}. In future climate studies, RCMs are often used to dynamically downscale precipitation to a local scale. However, even though the relevance of RCMs' outputs increases, a bias correction of the outputs is often still required, particularly in complex terrain. Moreover, their application is computer-intensive, which makes it difficult to cover all combinations of climate scenarios and GCMs. Therefore, the idea is to bypass the small-scale simulations and to go from the large-scale situation to the end variables such as precipitation by statistical downscaling \citep{Maraun2010}.

@@

The method can be designed with multiple successive subsampling steps, or analogy levels, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criterion itself, and finally, the number of analogue situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep{Bontron2004, Radanovics2013}, i.e. by optimizing each single parameter, one at a time, in an arbitrarily chosen order, with no or little reassessment. This sequential approach therefore has strong limitations: (i) it cannot automatically choose the optimal pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it cannot easily handle new degrees of freedom, such as a possible weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictor combinations (variables, pressure levels, temporal windows), and presents a high risk of ending in a local optimum because of subjective initial choices and lack of consideration of parameter inter-dependencies. Other calibration methods exist for specific applications, such as radar images \citep{Panziera2011, Foresti2015a}.

Aiming to overcome these limitations, a global optimization by genetic algorithms (GAs) was introduced. An intensive assessment resulted in recommendations to parametrize GAs in order to optimize AMs successfully \citep{Horton2016}. The present paper is based on these recommendations, and applies them to precipitation prediction for the upper Rh\^{o}ne catchment in the Swiss Alps, using AMs of varying complexity. It aims at illustrating the relevance of a fully automatic, objective, and global, optimization technique for AMs. The applications are indeed numerous, as AMs have to be adapted to every new location they are applied, or to any new predictand they should predict.

The data, AMs, and optimization techniques (sequential and GAs) are presented in Section \ref{sec:data_methods}. The results are first given for the optimization of the analogy of atmospheric circulation only (Section \ref{sec:optim_circul}), before being extended to a method adding a second level of analogy on moisture variables (Section \ref{sec:optim_moisture}). General discussions (Section \ref{sec:discussion}) and conclusions (Section \ref{sec:conclusions}) follow.

@@

The study area is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l.\ and the area is 5524~km$^{2}$. This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project aimed at real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. Even though the region is rather small, the meteorological influences related to extreme weather conditions vary substantially within it \citep[see][]{Horton2012}. Indeed, a high spatial variability of precipitation climatology exists, which is due to the complex orography of the region, and the mix of various meteorological influences. Based on different climatological analyses, the precipitation gauge stations in the catchment were clustered in ten subregions (Fig.\ \ref{fig:map}):

@@

AMs rely on two types of data: predictors, which are atmospheric variables describing the state of the atmosphere at a synoptic scale, and the predictand, which is the local weather variable one wants to predict.

@@

The predictand (which is to be predicted) is here the daily precipitation (6~a.m. to 6~a.m. the next day) measured at the MeteoSwiss network stations, for the period 1961--2008 . The time series from every available gauge station were averaged over the ten subregions (Fig.\ \ref{fig:map}), which were approximately 500~km$^{2}$ each, in order to smooth local effects \citep{Obled2002, Marty2012}. This helps account for local variability, mainly when convective processes are involved, which slightly increases the prediction skill.

It must be stressed that the predictand here is a temporally cumulated variable, compared to the meteorological predictors, which may be considered instantaneous. Depending on the duration of the accumulation period (here 24~h, but could have been 6~h, 12~h, or more than 24~h), the choice of predictors will vary.

The 48-yr precipitation dataset was divided into a calibration period (CP) and a validation period (VP). Using data independent of the CP to validate the results is very important in order to assess the robustness of the proposed improvements and to avoid over-parametrization of the method.

In order to reduce potential biases related to trends linked to climate change or to the evolution in measurement techniques, the selection of the VP was evenly distributed over the entire series \citep{BenDaoud2010}. Thus, one out of every six years was selected for validation, which represents a total of 8 years for the VP and 40 for the CP. This choice of sequence was made in order to have similar statistical characteristics between the CP and VP.

\subsection{Analogue method}
\label{sec:references}

Multiple variations of the analogue method exist, most of which are not detailed here \cite[see][for a more comprehensive listing]{BenDaoud2016}. However, there are mainly two parameterizations that are most often used for precipitation prediction and that are considered as reference: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

The method based on the analogy of synoptic circulation consists of the following steps (Table \ref{table:params_R1}): the similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criterion \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window:

\begin{equation}
\label{eq:S1}
S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the difference in geopotential height between the \textit{i}-th pair of adjacent points of gridded data describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both North and East directions over the selected spatial domain. The smaller the S1 values, the more similar the pressure fields.

\citet{Bontron2005} show that the geopotential height at 500~hPa (Z500) and 1000~hPa (Z1000) are the best first predictors of the NCEP/NCAR reanalysis I dataset, and that the S1 criterion performs better than scores based on absolute distances. The reason for such better results is that the S1 criterion allows comparison of the circulation patterns, by means of the gradients, rather than the absolute value of the geopotential height, which better represent the flow direction. To cope with seasonal effects, candidate dates are extracted within a period of four months centred around the target date, for every year of the archive. This method using two geopotential heights is named 2Z.

The $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day. The number of analogues, $N_{1}$, is a parameter to calibrate. Then, the daily observed precipitation amount for the $N_{1}$ resulting dates provide the empirical conditional distribution, considered as the probabilistic prediction for the target day.

The other most well-known parametrization adds a second level of analogy on the moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for France is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirms that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis I considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled within the $N_{1}$ analogues of the atmospheric circulation, to end up with a smaller number of analogue situations. When this second level of analogy is added, a higher number of analogues $N_{1}$ is kept on the first level. Moisture fields are not as well-predicted by NWP models as pressure variables. This implies that the 2Z-2MI method, when used in real-time forecasting, is very dependent on the skill of the NWP model in predicting moisture fields, and thus its use is often restricted to the first lead times.

@@

The performance assessment in the present context consists of verifying the prediction of an ensemble probabilistic technique. The set of precipitation values collected with each analogue can be considered as a sample drawn from the conditional distribution associated with the current circulation. The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example, of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $n$ can be written as:

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{n} \sum_{i=1}^{n} \left(  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right) 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value 1 otherwise. The mean CRPS is averaged on the calibration, respectively, the validation periods, on all days .

In order to compare the value of the score relative to a reference, one often considers its skill score expression, and uses the climatological distribution of precipitation from the entire archive as the reference. The CRPSS (Continuous Ranked Probability Skill Score) is thus defined as follows:

@@

Note, however, that the choice of reference does not matter so much when assessing potential improvements of the method, since we consider more its relative increase or decrease rather than the CRPSS absolute value.

@@

AMs are usually calibrated by a semi-automatic sequential procedure, as elaborated by \citet{Bontron2004} \cite[see also ][]{Radanovics2013, BenDaoud2016}. The calibration technique optimizes the spatial windows in which the predictors are compared and the number of analogues for every level of analogy, by maximizing the performance score (CRPSS). However, the different analogy levels are calibrated sequentially, and the meteorological variables, pressure levels, and temporal windows are chosen manually. The procedure, as defined by \citet{Bontron2004}, consists of the following steps:

\begin{enumerate}
	\item Manual selection of the following parameters:
	\begin{enumerate}
		\item Meteorological variable
		\item Pressure level
		\item Temporal window (hour of the day)
		\item Initial analogue numbers
	\end{enumerate}

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification, for the analogy level considered, of the most skilled unitary cell of all predictors jointly, over a large domain, by a full scanning of the grid.
		\item From this most skilled cell, the spatial window is expanded by successive iterations in the direction of greater performance gain until no improvement is reached.
		\item The number of analogue situations $N_{1}$ is then reconsidered and optimized for the current level of analogy.
	\end{enumerate}
	\item A new level of analogy can then be added, based on other variables (such as the moisture index) with the chosen pressure levels, temporal windows, and initial number of analogues $N_{2}$. The procedure starts again from step 2 (calibration of the spatial window and the number of analogues) for the new level. The parameters calibrated on the previous analogy levels are fixed and do not change. 
	\item Finally, the numbers of analogues $N_{1}$ and $N_{2}$ for the different levels of analogy are reassessed by systematic increments.
\end{enumerate}

@@

Genetic algorithms (GAs) were developed by \citet{Holland1992b} and \citet{Goldberg1989}. They are part of the Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, which were inspired by some mechanisms in biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs seek the global optimum on a complex surface, theoretically without restriction, which is of interest for AMs, which are characterized by a complex high-dimensional error function having multiple local optima. Practically, GAs allow rapidly approaching satisfactory solutions, but they are not guaranteed to provide the optimum solution \citep{Zitzler2004a}. It is indeed mainly a matter of time. When the optimizer gets closer to the global optimum, any new improvement takes more time to appear (see for example the slow-down of the improvements over generations in Figure \ref{fig:evolution}), and the final adjustment of the parameters can be very time consuming \citep{Back1993a}. For problems that require a significant amount of time to evaluate the objective function, as in the case of AMs (because it needs to make a prediction for every day of the CP), the number of generations has to be limited in order to ensure a reasonable processing time. Thus, different acceptable solutions can result from one or more optimization runs \citep{Holland1992b}. This is both a strength and a weakness of GAs: they are very good at exploring complex parameter spaces in order to identify the most promising areas, but they will not necessarily always find the best solution with the optimal values for all parameters \citep{Holland1992b}.

The optimizations here were performed based on the recommended GA parametrization for AMs as described in \citet{Horton2016}. As the optimization is mostly sensitive to the mutation operator (that randomly changes some values in the parameter sets), parallel optimizations are considered with variants of this operator, according to \citet{Horton2016}:

\begin{itemize}
	\item 3x non-uniform mutation \citep{Michalewicz1996} with varying parameters
	\item 1x multi-scale mutation \citep{Horton2016}
	\item 2x chromosome of adaptive search radius \citep{Horton2016}.
\end{itemize}

A population size of 500 individuals (i.e. parameter sets of the AM to be detailed below) was considered, and the optimization was stopped when the best individual (with the highest CRPSS performance score) did not evolve for 20 generations (cycles of optimization).

\section{Optimization of circulation analogy}
\label{sec:optim_circul}

The analogy of the atmospheric circulation was optimized for the ten subregions (Section \ref{sec:case_study}) independently. We started from the simplest AM, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. The detailed results of the intermediate stages are not provided in this paper \citep[see][for the details]{Horton2012a}.

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose simultaneously the number of analogues, both spatial windows with no overlapping constraint (i.e. they can differ from one pressure level to another), as well as the temporal windows (hours of observation of the geopotential), which cannot be achieved with the sequential calibration technique. The performance score (CRPSS) was slightly improved, with these limited degrees of freedom, relative to the 2Z reference method calibrated with the sequential procedure. Some tests showed that most of the gains were due to the non-overlapping spatial windows. This demonstrated that the optimizer was able to obtain relevant parameters for a simple method.
